---
title: "In-class_Ex09"
author: "Enrico Limberg"
---

## **1.0 Load R Packages**

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts)
```

## **2.0 Import the Data**

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
```

```{r}
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```

# 3.0 Data Wrangling

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>%
  mutate(Weekday = wday(SentDate,
                        label = TRUE,
                        abbr = FALSE))
```

The code chunk below is to aggregate the data based on the source, target and the weekday to simplify the df. The code [***filter(source!=target)***]{.underline} is to avoid same source and target value.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

The code below will show how many times each node [**are linked**]{.underline} to each other.

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

Compute the network metrics [**betweenness, closeness, centrality measure, etc**]{.underline} to interpret the meaning of the network graph.

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```
