---
title: "Hands-on_Ex10"
author: "Enrico Limberg"
date: "March 26, 2025"
---

# **1.0 Information Dashboard Design: R methods**

## **1.1 Overview**

By the end of this hands-on exercise, we will be able to:

-   create ***bullet chart*** by using [**ggplot2**]{.underline},

-   create ***sparklines*** by using [**ggplot2**]{.underline} ,

-   build industry standard dashboard by using [**R Shiny**]{.underline}.

## **1.2 Getting started**

For the purpose of this hands-on exercise, the following R packages will be used.

```{r}
pacman::p_load(lubridate, ggthemes, reactable,
reactablefmtr, gt, gtExtras, tidyverse)
```

## **1.3 Importing Microsoft Access database**

### **1.3.1 The data set**

For the purpose of this study, a personal database in Microsoft Access mdb format called *Coffee Chain* will be used.

### **31.3.2 Importing database into R**

In the code chunk below, [***odbcConnectAccess()***]{.underline} of [**RODBC**]{.underline} package is used used to import a database query table into R.

```{r}
library(RODBC)
con <- odbcConnectAccess2007('data/Coffee Chain.mdb')
coffeechain <- sqlFetch(con, 'CoffeeChain Query')
write_rds(coffeechain, "data/CoffeeChain.rds")
odbcClose(con)
```

### **1.3.3 Data Preparation**

The code chunk below is used to aggregate ***Sales*** and ***Budgeted Sales*** at the Product level.

```{r}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

### **1.3.4 Bullet chart in ggplot2**

The code chunk below is used to plot the ***bullet charts*** using ggplot2 functions.

```{r}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

## **1.4 Plotting sparklines using ggplot2**

In this section, we will learn how to plot [***sparklines***]{.underline} by using ggplot2.

### **1.4.1 Preparing the data**

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

The code chunk below is used to compute the [***minimum, maximum and end***]{.underline} othe the month sales.

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

The code chunk below is used to compute the [***25 and 75 quantiles.***]{.underline}

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)
```

### **1.4.2 sparklines in ggplot2**

```{r, fig.height=10,fig.width=10}
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

## **1.5 Static Information Dashboard Design: gt and gtExtras methods**

In this section, we will learn how to create static information dashboard by using [**gt**]{.underline} and [**gtExtras**]{.underline} packages.

### **1.5.1 Plotting a simple bullet chart**

In this section, we will learn how to prepare a ***bullet chart report*** by using functions of [**gt**]{.underline} and [**gtExtras**]{.underline} packages.

```{r}
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## **1.6 sparklines: gtExtras method**

Before we can prepare the sales report by product by using gtExtras functions, code chunk below will be used to prepare the data.

```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()
```

::: callout-important
## Important note

To use the [**gtExtras**]{.underline} functions, we need to pass data.frame in **list columns**, thus using the code chunk below we will convert the report data.frame into list columns.
:::

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

### **1.6.1 Plotting Coffechain Sales report**

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)
```

### **1.6.2 Adding statistics**

**First**, calculate summary statistics by using the code chunk below.

```{r}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

### **1.6.3 Combining the data.frame**

**Next,** use the code chunk below to add the statistics on the table.

```{r}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

```{r}
sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )
```

```{r}
sales_data = left_join(sales, spark)
```

### **1.6.4 Plotting the updated data.table**

```{r,fig.width=10}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```

### **1.6.5 Combining bullet chart and sparklines**

We can combining the bullet chart and sparklines through the steps below.

```{r}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 
```

```{r}
sales_data = sales_data %>%
  left_join(bullet)
```

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## **1.7 Interactive Information Dashboard Design: [reactable and reactablefmtr methods]{.underline}**

In this section, we will learn how to create interactive information dashboard by using [**reactable**]{.underline} and [**reactablefmtr**]{.underline} packages.

To build an ***interactive sparklines***, we need to install and load [**dataui**]{.underline} R package.

```{r}
remotes::install_github("timelyportfolio/dataui")
```

```{r}
library(dataui)
```

### **1.7.1 Plotting interactive sparklines**

Similar to ***gtExtras***, to plot an interactive sparklines by using ***reactablefmtr*** package we need to prepare the list field.

```{r}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

Next, ***react_sparkline*** will be used to plot the sparklines.

```{r}
reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

### **1.7.2 Changing the pagesize**

[**By default**]{.underline} the pagesize is 10. With the code chunk below, arguments [**defaultPageSize**]{.underline} is used to change the default setting.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

### **1.7.3 Adding points and labels**

[***highlight_points***]{.underline} argument is used to show the ***minimum and maximum values*** points and [***label***]{.underline} argument is used to label ***first and last values.***

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```

### **1.7.4 Adding reference line**

[***statline***]{.underline} argument is used to show the [**mean line.**]{.underline}

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```

### **1.7.5 Adding bandline**

Beside reference line, [**bandline**]{.underline} can be added by using the [***bandline***]{.underline} argument.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```

### **1.7.6 Changing from sparkline to sparkbar**

Instead of displaying the values as **sparklines**, we can display them as **sparkbars** as shown below.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```
