---
title: "Hands-on_Ex05"
author: "Enrico Limberg"
date: "February 8, 2025"
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **9 Creating Ternary Plot with R**

## **9.1 Overview**

Ternary plots are a way of displaying the distribution and variability of three-part compositional data. (For example, the proportion of aged, economy active and young population or sand, silt, and clay in soil.) It’s display is a triangle with sides scaled from 0 to 1. Each side represents one of the three components. A point is plotted so that a line drawn perpendicular from the point to each leg of the triangle intersect at the component values of the point.

In this hands-on, you will learn how to build ternary plot programmatically using R for visualising and analysing population structure of Singapore.

The hands-on exercise consists of four steps:

-   Install and launch **tidyverse** and **ggtern** packages.

-   Derive three new measures using *mutate()* function of **dplyr** package.

-   Build a static ternary plot using *ggtern()* function of **ggtern** package.

-   Build an interactive ternary plot using *plot-ly()* function of **Plotly R** package.

## **9.2 Installing and launching R packages**

For this exercise, two main R packages will be used in this hands-on exercise, they are:

-   [***ggtern***]{.underline}**,** a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots.

-   [***Plotly R,***]{.underline} an R package for creating interactive web-based graphs via plotly’s JavaScript graphing library, plotly.js . The **plotly R** libary contains the *ggplotly* function, which will convert **ggplot2** figures into a Plotly object.

We will also need to ensure that selected **tidyverse** family packages namely: **readr**, **dplyr** and **tidyr** are also installed and loaded.

In this exercise, version 3.2.1 of **ggplot2** will be installed instead of the latest version of **ggplot2**. This is because the current version of **ggtern** package is not compatible to the latest version of **ggplot2**.

The code chunks below will accomplish the task.

```{r}
pacman::p_load(plotly, ggtern, tidyverse)
```

## **9.3 Data Preparation**

### **9.3.1 The data**

For the purpose of this hands-on exercise, the [**Singapore Residents by Planning AreaSubzone, Age Group, Sex and Type of Dwelling, June 2000-2018**]{.underline} data will be used. The data set has been downloaded and included in the data sub-folder of the hands-on exercise folder. It is called *respopagsex2000to2018_tidy.csv* and is in csv file format.

### **9.3.2 Importing Data**

To important *respopagsex2000to2018_tidy.csv* into R, ***read_csv()*** function of **readr** package will be used.

```{r}
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

### **9.3.3 Preparing the Data**

Next, use the ***mutate()*** function of **dplyr** package to derive three new measures, namely: young, active, and old.

```{r}
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)
```

## **9.4 Plotting Ternary Diagram with R**

### **9.4.1 4.1 Plotting a static ternary diagram**

Use ***ggtern()*** function of **ggtern** package to create a simple ternary plot.

```{r}
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```
